Formalizing Stack Safety as a Security Property

### 2021-09-29 submission4/main.tex
Moved testing to the end

Observable properties allow us to defer enforcement until a property
violation would become visible. Now we can identify when lazy policies
miss such violations, which are again susceptible to testing;
Table~\ref{tab:lazy-bug-table} summarizes the results of our evaluation.
The Lazy Tagging and Clearing
micro-policy from Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} in fact
allows violations of observable stack safety!
Our implementation of this policy, {\em PER\_DEPTH\_TAG}, has an error that we find
in testing. If, at depth \(d\), a callee \(A\) writes into its caller's
frame, then returns, and later another callee \(B\) at depth \(d\) reads the same
address and prints it, \(A\) has violated integrity in a way that later becomes visible,
and \(B\) has likewise violated confidentiality!

Lazy Tagging and Clearing can be repaired if, instead of tagging each function
activation with its depth in the stack, we generate a fresh activation
identifier on each call, which prevents a program from exploiting stale
tags from previous activations.
(A related mechanism was explored in the
Static Authorities policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18}, which
associates a unique \emph{static} identifier to each function, which is shared
by all activations---but that policy allows leaks between
instances of the same function.)

We test {\em Lazy Per-activation Tagging and Clearing}, in which
each activation is tagged uniquely, and find that it obeys stepwise
confidentiality and observational integrity. To round out our mutations testing
we also check an equivalent to {\em LOAD\_NO\_CHECK} and a version where
stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
taking twice as long to catch as the next longest.

\begin{table}[]
\centering
\begin{tabular}{c|c|c|c}
        Bug & Property Violated & MTTF (s) & Tests \\
  \hline
        {\em PER\_DEPTH\_TAG} & Obs. Integrity & 189.7 & 8342.5  \\
        {\em LOAD\_NO\_CHECK}  & Obs. Integrity & 23.5 & 12.0 \\
        {\em LOAD\_NO\_CHECK}  & Confidentiality & 19.2 & 695.5 \\
        {\em STORE\_NO\_UPDATE} & Obs. Integrity & 70 & 80.6  \\
        {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\
\end{tabular}
\vspace*{1em}
\caption{MTTF for finding bugs in erroneous lazy policy enforcement mechanisms}
\vspace*{-2em}
\label{tab:lazy-bug-table}
\end{table}

@@

void f(int &a) {
  if(!stash)
    stash = &a;
  *stash = 5;
  return;
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}[t]{.47\textwidth}
{\small
\begin{verbatim}
   stash: .word 0
1  main:  add $2,%sp      ; alloc frame
2         mov %sp,%ra
3         add -2,%ra      ; ra = &x
4         call f          ; SHARE: %sp-2
5         mov $0,-2(%sp)  ; x = 0
6         mov %sp,%ra
7         add -1,%ra      ; ra = &y
8         call f          ; SHARE: %sp-1
9         mov -2(%sp),%ra ; set return
10        sub $2,%sp      ; dealloc frame
11        ret

20 f:     cmp stash, $0   ; stash = 0?
21        breq #13        ; branch if so
22        mov %ra,stash   ; stash = &a
23        jmp #15
24        mov $5, (stash) ; *stash = 5
25        ret
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}[t]{.27\textwidth}
    \begin{center}
    \begin{tabular}{l l}
      {\tt 20} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{\passdesc{0}{1}}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{\sealdesc{0}}}]%
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 25} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}
      \memory{1}{\mainsealc}
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{5}
      \\
      {\tt 8} &
      \memoryaddrs[(0)]{12em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 20} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainsealc}[{\makebox[0pt]{\sealdesc{0}}}]%
      \memory{1}{\mainpassc}[{\makebox[0pt]{\passdesc{0}{1}}}]%
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 25} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{\bf 5}
\end{tabular}
\end{center}
\end{subfigure}
\caption{A violation with pass-by-reference}
\label{fig:passing}
\end{figure*}

@@

We validate the micro-policy of the previous section by
systematically testing that it satisfies eager stack integrity and
confidentiality. We use a Coq specification of the RISC-V
architecture\bcp{change from footnote to citation :-)}\footnote{Available at
  \url{https://github.com/mit-plv/riscv-coq}.}
%\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since the MIT people did it}}
and extend it with a
runtime monitor implementing the stack-safety
micro-policy. We
  chose the Coq proof assistant as the setting for our implementation
  to ensure that our coinductive trace
  definitions are well-formed, and reason about them, and to leverage the power of the
  QuickChick property-based testing framework~\cite{Pierce:SF4}.

@@

%\paragraph*{Generators}
%
%To write such generators we build on the work of
%Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
%introduced {\em generation by execution} to produce progams that lead
%machines towards interesting behaviors. Generation by
%execution receives as an input a partially instantiated machine state
%and attempts to generate an instruction (or a sequence of instructions
%such as a blessed sequence) that makes sense locally (e.g., jumps go
%to a potentially valid code location and loads read from a
%potentially valid stack location). Then we step the machine and repeat
%the process until we generate or execute some target number of
%instructions, or reach a point where the machine cannot step
%any more.
%
%We extend this technique to keep track of the control flow behavior of
%the program being generated: each time a call or return sequence is
%generated, we ensure that the appropriate policy tags and code
%annotations are set for the entry or return points.% At the same time,
%%we allow the generation to sometimes relax those constraints,
%%introducing potentially ill-formed flows: this causes our programs
%%to failstop when executed in conjunction with the policy monitor,
%%but also allows for revealing errors in our setup.\apt{??}
%
%In addition, we need to further extend this technique to handle the nested
%nature of our confidentiality property: rather than just generating two
%initial machines that are variants of one another and letting them
%execute to test for noninterference, we generate a new variant
%{\em every time a call is made} and check confidentiality for the
%subtrace produced from that variant state until its corresponding
%return. As a result, a ``single'' confidentiality test compactly
%checks the confidentiality of multiple nested calls.

\paragraph*{Findings}

@@

\begin{table}[]
\centering
\begin{tabular}{c|c|c|c}
  Bug & Property Violated & MTTF (s) & Tests \\
  \hline
      {\em LOAD\_NO\_CHECK}  & Confidentiality & 24.2 & 13.3 \\
      {\em STORE\_NO\_CHECK} & Integrity & 26.9 & 26 \\
      {\em HEADER\_NO\_INIT} & Integrity & 69.5 & 76.3 \\
\end{tabular}
\vspace*{1em}
\caption{MTTF for finding bugs in erroneous eager policy enforcement mechanisms}
\vspace*{-2em}
\label{tab:bug-table}
\end{table}

The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
Table~\ref{tab:bug-table}, along with the average number of tests
it took to find the failure. Experiments were run in a desktop
machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.
% \leo{Something like: There is really no baseline to compare this
  %against, but include it here for the future/reasons?}
%
The artificially injected bugs weaken the tag propagation
and checking mechanism.
%
For example, when loading from a stack location, the correct
micro-policy needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
that restriction by not checking the depth equality (row {\em
  LOAD\_NO\_CHECK}).
%
Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing that constraint causes violations to the integrity
property (row {\em STORE\_NO\_CHECK}).

@@

>>>>

Observable properties allow us to defer enforcement until a property
violation would become visible. Now we can identify when lazy policies
miss such violations. It turns out that {\em Lazy Tagging and Clearing},
as written, falls short of observational stack safety.
If, at depth \(d\), a callee \(A\) writes into its caller's
frame, then returns, and later another callee \(B\) at depth \(d\) reads the same
address and prints it, \(A\) has violated integrity in a way that later becomes visible,
and \(B\) has likewise violated confidentiality!

{\em Lazy Tagging and Clearing} can be repaired if, instead of tagging each function
activation with its depth in the stack, we generate a fresh activation
identifier on each call, which prevents a program from exploiting stale
tags from previous activations.
(A related mechanism was explored in the
Static Authorities policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18}, which
associates a unique \emph{static} identifier to each function, which is shared
by all activations---but that policy allows leaks between
instances of the same function.) We validate this repaired micro-policy in
\ref{sec:testing}.

@@

void f(int &a) {
  if(!stash)
    stash = &a;
  *stash = 5;
  return;
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}[t]{.4\textwidth}
{\small
\begin{verbatim}
   stash: .word 0
1  main:  add $2,%sp      ; alloc frame
2         mov %sp,%ra
3         add -2,%ra      ; ra = &x
4         call f          ; SHARE: %sp-2
5         mov $0,-2(%sp)  ; x = 0
6         mov %sp,%ra
7         add -1,%ra      ; ra = &y
8         call f          ; SHARE: %sp-1
9         mov -2(%sp),%ra ; set return
10        sub $2,%sp      ; dealloc frame
11        ret

20 f:     cmp stash, $0   ; stash = 0?
21        breq #13        ; branch if so
22        mov %ra,stash   ; stash = &a
23        jmp #15
24        mov $5, (stash) ; *stash = 5
25        ret
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}[t]{.25\textwidth}
    \begin{center}
    \begin{tabular}{l l}
      {\tt 20} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{\passdesc{0}{1}}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{\sealdesc{0}}}]%
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 25} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}
      \memory{1}{\mainsealc}
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{5}
      \\
      {\tt 8} &
      \memoryaddrs[(0)]{12em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 20} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainsealc}[{\makebox[0pt]{\sealdesc{0}}}]%
      \memory{1}{\mainpassc}[{\makebox[0pt]{\passdesc{0}{1}}}]%
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 25} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{\bf 5}
\end{tabular}
\end{center}
\end{subfigure}
\caption{A violation with pass-by-reference}
\label{fig:passing}
\end{figure*}

@@

We validate the micro-policy of the previous section by
systematically testing that it satisfies eager stack integrity and
confidentiality. We use a Coq specification of the RISC-V
architecture\bcp{change from footnote to citation :-)}\footnote{Available at
  \url{https://github.com/mit-plv/riscv-coq}.}
and extend it with a runtime monitor implementing a stack-safety
micro-policy. We chose the Coq proof assistant as the setting for our implementation
to ensure that our coinductive trace definitions are well-formed, and reason about
them, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.

@@

To write such generators we build on the work of
Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
introduced {\em generation by execution} to produce progams that lead
machines towards interesting behaviors. We extend the technique by tracking
control flow information so that programs will usually --- but not always ---
return to their call sites.

%\paragraph*{Generators}
%
% Generation by
%execution receives as an input a partially instantiated machine state
%and attempts to generate an instruction (or a sequence of instructions
%such as a blessed sequence) that makes sense locally (e.g., jumps go
%to a potentially valid code location and loads read from a
%potentially valid stack location). Then we step the machine and repeat
%the process until we generate or execute some target number of
%instructions, or reach a point where the machine cannot step
%any more.
%
%We extend this technique to keep track of the control flow behavior of
%the program being generated: each time a call or return sequence is
%generated, we ensure that the appropriate policy tags and code
%annotations are set for the entry or return points.% At the same time,
%%we allow the generation to sometimes relax those constraints,
%%introducing potentially ill-formed flows: this causes our programs
%%to failstop when executed in conjunction with the policy monitor,
%%but also allows for revealing errors in our setup.\apt{??}
%
We need to further extend typical testing to handle the nested
nature of confidentiality: rather than just generating two
initial machines that are variants of one another and letting them
execute to test for noninterference, we generate a new variant
{\em every time a call is made} and check confidentiality for the
subtrace produced from that variant state until its corresponding
return. As a result, a ``single'' confidentiality test compactly
checks the confidentiality of multiple nested calls.

Our primary testing targets are the eager {\em Depth Isolation}
and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.
We test the former against stepwise integrity and confidentiality, and
the latter against their observational counterparts.

@@

For example, when loading from a stack location, {\em Depth Isolation}
needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
that restriction by not checking the depth equality (row {\em
  LOAD\_NO\_CHECK\_DI}).

Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing that constraint causes violations to the integrity
property (row {\em STORE\_NO\_CHECK}).

\begin{table}[]
\centering
\begin{tabular}{c|c|c|c}
  Bug & Property Violated & MTTF (s) & Tests \\
  \hline
      {\em LOAD\_NO\_CHECK\_DI}  & Confidentiality & 24.2 & 13.3 \\
      {\em STORE\_NO\_CHECK} & Integrity & 26.9 & 26 \\
      {\em HEADER\_NO\_INIT} & Integrity & 69.5 & 76.3 \\
  \hline
  \hline
      {\em PER\_DEPTH\_TAG} & Obs. Integrity & 189.7 & 8342.5  \\
      {\em LOAD\_NO\_CHECK\_LT}  & Obs. Integrity & 23.5 & 12.0 \\
      {\em LOAD\_NO\_CHECK\_LT}  & Confidentiality & 19.2 & 695.5 \\
      {\em STORE\_NO\_UPDATE} & Obs. Integrity & 70 & 80.6  \\
      {\em STORE\_NO\_UPDATE} & Confidentiality & 4.9 & 88.5 \\
  \hline
\end{tabular}
\vspace*{1em}
\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
\vspace*{-2em}
\label{tab:bug-table}
\end{table}

The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
Table~\ref{tab:bug-table}, along with the average number of tests
it took to find the failure. Experiments were run in a desktop
machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.

@@

Table~\ref{tab:lazy-bug-table} summarizes the results of our evaluation on lazy micro-policies.
The {\em Lazy Tagging and Clearing} micro-policy implemented as {\em PER\_DEPTH\_TAG},
fails in testing, in cases where data is leaked between sequential calls.
To round out our mutations testing
we also check an equivalent to {\em LOAD\_NO\_CHECK\_DI}, {\em LOAD\_NO\_CHECK\_LT}
and a version where stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
taking twice as long to catch as the next longest.
